---
title: "Week 3 Assessment"
output: pdf_document
---
# Week 3 Assessment
#Exercise 1

```{r combined_graph}
library(tidyverse)
library(cowplot)
library(ggpubr)

nhanes <- read_csv("C:/Users/kayla/Downloads/cleaned_NHANES.csv")
diet <- read_csv("C:/Users/kayla/Downloads/diet.csv")

# Age distribution separated by gender
age_gender_graph <- ggplot(nhanes, aes(x = age)) +
  geom_histogram(binwidth = 5, boundary = 0) +
  facet_wrap(~ gender) +
  labs(
    title = "Age Distribution by Gender",
    x = "Age",
    y = "Count"
  )

# Ethnicity graph showing gender proportions
ethnicity_gender_graph <- ggplot(
  nhanes,
  aes(x = ethnicity_2, fill = gender)
) +
  geom_bar(position = "fill") +
  labs(
    title = "Gender Proportions by Ethnicity",
    x = "Ethnicity",
    y = "Proportion",
    fill = "Gender"
  )

age_gender_graph
ethnicity_gender_graph

# Combine using cowplot
combined_cowplot <- plot_grid(
  age_gender_graph,
  ethnicity_gender_graph,
  labels = c("A", "B"),
  ncol = 2
)

combined_cowplot

# Combine using ggpubr
combined_ggarrange <- ggarrange(
  age_gender_graph,
  ethnicity_gender_graph,
  labels = c("A", "B"),
  ncol = 2,
  nrow = 1
)

combined_ggarrange
```
#I prefer ggarrange() because the syntax is straightforward and makes it easy to arrange and label multiple plots in one figure.
#Exercise 2
```{r age graph}
# AGE GRAPH
age_graph <- ggplot(nhanes, aes(x = age)) +
  geom_histogram(binwidth = 5, boundary = 0) +
  labs(
    title = "Distribution of Age",
    x = "Age",
    y = "Count"
  )

age_graph

# GENDER GRAPH
gender_graph <- ggplot(nhanes, aes(x = gender)) +
  geom_bar() +
  labs(
    title = "Distribution of Gender",
    x = "Gender",
    y = "Count"
  )

gender_graph

# ETHNICITY 1 GRAPH
ethnicity1_graph <- ggplot(nhanes, aes(x = ethnicity_1)) +
  geom_bar() +
  labs(
    title = "Distribution of Ethnicity 1",
    x = "Ethnicity",
    y = "Count"
  )

ethnicity1_graph

# ETHNICITY 2 GRAPH
ethnicity2_graph <- ggplot(nhanes, aes(x = ethnicity_2)) +
  geom_bar() +
  labs(
    title = "Distribution of Ethnicity 2",
    x = "Ethnicity",
    y = "Count"
  )

ethnicity2_graph

# Missing values
sum(is.na(nhanes$age))
sum(is.na(nhanes$gender))
sum(is.na(nhanes$ethnicity_1))
sum(is.na(nhanes$ethnicity_2))

# Compare ethnicity categories
table(nhanes$ethnicity_1)
table(nhanes$ethnicity_2)

# Save graphs
ggsave("age_graph.png", plot = age_graph)
ggsave("gender_graph.png", plot = gender_graph)
ggsave("ethnicity1_graph.png", plot = ethnicity1_graph)
ggsave("ethnicity2_graph.png", plot = ethnicity2_graph)

# Remove ethnicity_1
nhanes <- nhanes %>%
  select(-ethnicity_1)

names(nhanes)
```

#I chose to keep ethnicity_2 because it provides more detailed ethnicity categories by identifying Asian participants separately. ethnicity_1 groups those participants into the broader “Other” category, so ethnicity_2 preserves more information.



#Exercise 3
```{r diet_change}

diet_change <- diet %>%
  group_by(Participant) %>%
  mutate(
    starting_weight = Weight[match(0, Week)],
    weight_change = Weight - starting_weight
  ) %>%
  ungroup()

weight_change_graph <- ggplot(
  diet_change,
  aes(x = Week, y = weight_change, group = Participant)
) +
  geom_line() +
  geom_point() +
  labs(
    title = "Weight Change Over Time",
    x = "Week",
    y = "Change From Starting Weight"
  )

weight_change_graph
```
#This graph is easier to understand because it shows how much each participant’s weight changed from their own starting weight rather than comparing their actual body weights. This makes the direction and amount of weight change easier to see.




#Exercise 4

```{r bp_data}

bp_data <- nhanes %>%
  mutate(
    avg_systolic_bp = rowMeans(
      across(c(systolic_bp_1, systolic_bp_2, systolic_bp_3, systolic_bp_4)),
      na.rm = TRUE
    ),
    avg_diastolic_bp = rowMeans(
      across(c(diastolic_bp_1, diastolic_bp_2, diastolic_bp_3, diastolic_bp_4)),
      na.rm = TRUE
    ),
    avg_systolic_bp = ifelse(is.nan(avg_systolic_bp), NA, avg_systolic_bp),
    avg_diastolic_bp = ifelse(is.nan(avg_diastolic_bp), NA, avg_diastolic_bp)
  ) %>%
  select(
    patient_id,
    age,
    gender,
    avg_systolic_bp,
    avg_diastolic_bp
  )

head(bp_data)
